---
title: "Auto Box Plots"
author: "John Quensen"
date: "5/20/2020"
output:
  pdf_document: default
  html_document: default
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

An exercise in writing functions to automate plotting.  

Load libraries and import data.  


```{r, tidy=TRUE, warning=FALSE, message=FALSE, echo=FALSE}
suppressPackageStartupMessages(library(phyloseq))
suppressPackageStartupMessages(library(QsRutils))
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(ggordiplots))
suppressPackageStartupMessages(library(hablar))

load("D:/Users/johnq/OneDrive/Kize/phyloseq/expt_sub.rda")
```

First, we have to subset our data to the treatment combinations we want to compare. We begin by sub-setting to harvest time and bulk soil.  

```{r}
bulk.har <- subset_samples(expt.sub, Sampling_time == "Harvest")
bulk.har <- subset_samples(bulk.har, Soil == "Bulk")
bulk.har <- prune_taxa(taxa_sums(bulk.har)>0, bulk.har)
bulk.har
```

Make a data tibble for plotting $\alpha$-diversities for the different cropping methods.   

```{r}
yr <- vegansam(bulk.har) %>% 
  rownames_to_column("Sample") %>% 
  select(Sample, Season, Crop) %>% 
  as_tibble()

temp <-   estimate_richness(bulk.har, measures = c("Observed", "Chao1", "Shannon")) %>% 
  rownames_to_column("Sample") %>% 
  select(-se.chao1) %>% 
  mutate(Pielou = Shannon/log(Observed)) %>% 
  as_tibble()

alpha.div <- 
  full_join(yr, temp, by = "Sample") %>% 
  select(-Sample) %>% 
  pivot_longer(c(Observed, Chao1, Shannon, Pielou), names_to = "Index") %>% 
  dplyr::rename(Diversity = value)
  
alpha.div
```

Subset to one season.  

```{r}
alpha.sub <- alpha.div %>% 
  hablar::convert(chr(Season)) %>% 
  dplyr::filter(Season == "2018-2019") %>% 
  droplevels("Season")
alpha.sub
unique(alpha.sub$Season)
```

Write a function to make a tibble giving coordinates of the compact letter displays for a box plot. Determine the compact letter displays using pairwise Wilcox tests.  

```{r}
df_ltrs <- function(df = df, grp = "Crop", value = "Diversity", alpha = 0.05) {
  rslt <- pairwise.wilcox.test(pull(df, Diversity), pull(df, Crop))
  ltrs <- make_letter_assignments(rslt, significance = alpha)$Letters
  max.val <- df %>% 
    dplyr::group_by_at(grp) %>% 
    dplyr::summarize_at(.vars = value, .funs = max)
  df.ltrs <- data.frame(max.val, ltrs)
  colnames(df.ltrs) <- c("x", "y", "ltrs")
  return(df.ltrs)
}
```

Write a function to make box plots with compact letter displays indicating significant differences among treatments.  

```{r}
boxplot_cld <- function(df, x, y, alpha) {
  plt <- ggplot(data = df, aes(x=Crop, y=Diversity)) + geom_boxplot()
  df.ltrs <- df_ltrs(df = df, grp = x, value = y, alpha = alpha)
  lmts <- get_plot_limits(plt)
  y.range <- lmts$ymax - lmts$ymin
  y.nudge <- 0.05 * y.range
  plt + geom_text(data = df.ltrs, aes(x=x, y=y, label=ltrs), nudge_y = y.nudge, size = 5)
}
```

Use the functions within a loop to make separate plots for each diversity index.  

```{r}
indices <- unique(alpha.sub$Index)
indices

for (index in indices) {
  df <- dplyr::filter(alpha.sub, Index == index)
  plt <- boxplot_cld(df=df, x="Crop", y="Diversity", alpha = 0.1)
  print(plt + ylab(index) + ggtitle("2018-2019"))
}
```

Here is another way to execute the plotting from within a loop. I do not execute it in this document, but try it yourself and see.  

```{r, eval=FALSE}
while (length(indices < 0)) {
  df <- dplyr::filter(alpha.sub, Index == index)
  plt <- boxplot_cld(df=df, x="Crop", y="Diversity", alpha = 0.1)
  print(plt + ylab(index) + ggtitle("2018-2019"))
  indices <- indices[-1]
}
```

You can also subset alpha.div to alpha.sub for other seasons and re-do the above, try different values of alpha, etc.  
