---
title: "Blooming Time"
author: "John Quensen"
date: "5/18/2020"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

2. What is the differencee and abundance of taxa comparing rhizosphere and bulk soil?

```{r, tidy=TRUE, warning=FALSE, message=FALSE, echo=FALSE}
suppressPackageStartupMessages(library(phyloseq))
suppressPackageStartupMessages(library(QsRutils))
suppressPackageStartupMessages(library(ade4))
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(ggordiplots))
suppressPackageStartupMessages(library(hablar))
suppressPackageStartupMessages(library(GUniFrac))
suppressPackageStartupMessages(library(RVAideMemoire))
pretty_int <- function(x) {
  formatC(x, format = "f", big.mark = ",", digits = 0)
}
load("D:/Users/johnq/OneDrive/Kize/phyloseq/expt_sub.rda")
```

```{r}
bloom <- subset_samples(expt.sub, Sampling_time == "Blooming")
bloom <- prune_taxa(taxa_sums(bloom)>0, bloom)
sample_variables(bloom)

# Tally the number of samples per treatment group.
num.trts <- vegansam(bloom) %>% 
  group_by(Season, Soil, Crop, Block) %>% 
  summarize(Count = n())
as.data.frame(num.trts) # Because otherwise all rows will not dispay.
```

I will ignore blocks for now and make an ordination of the samples based on weighted unifrac distances.  

```{r}
my.tree <- phy_tree(bloom)
my.otu <- veganotu(bloom)
unifracs <- GUniFrac(my.otu, my.tree,  alpha = c(0, 0.5, 1))
dw <- as.dist(unifracs$unifracs[, , "d_1"])
add <- !(is.euclid(dw))
sam <- vegansam(bloom)
grps <- paste(sam$Soil, sam$Crop, sep = "_")
ord <- cmdscale(dw, k=nsamples(bloom)-1, eig = TRUE, add = add)
gg_ordiplot(ord, groups = grps)
```

The bulk soil samples cluster together. If anything, the corn samples are slightly different from the others.  

The rhizosphere samples show more dispersion.    

Make another ordination for just the rhizosphere samples and see if there are signficant differences among crops.  

```{r}
blm.rhiz <- subset_samples(bloom, Soil == "Rhizosphere")
blm.rhiz <- prune_taxa(taxa_sums(blm.rhiz)>0, blm.rhiz)
dw.sub <- subset_dist(blm.rhiz, dw)
add <- !(is.euclid(dw.sub))
grps <- vegansam(blm.rhiz)$Crop
ord <- cmdscale(dw.sub, k=nsamples(blm.rhiz)-1, eig = TRUE, add = add)
gg_ordiplot(ord, groups = grps)
gg_ordiplot(ord, groups = grps, choices = c(1,3))
```

Tests for differences in dispersion.  

```{r}
rslt <- betadisper(dw.sub, group = grps, add = add)
rslt
permutest(rslt)
```

So dispersions are not significantly different. Proceed with PERMANOVA.  

```{r}
rslt <- adonis(dw.sub ~ grps)
rslt
```

Look for pairwise differences.  

```{r}
rslt <- pairwise.perm.manova(dw.sub, fact = grps)
rslt
```

All crops are significantly different. Find families signficantly different among the crops.    

```{r}
blm.rhiz <- tax_glom(blm.rhiz, taxrank = "Family")
blm.rhiz
library(RDPutils)
# It is necessacry to rename the taxa; otherwise R will
# rename any beginning with a digit when the indval
# report is made.
taxa_names(blm.rhiz) <- make_otu_names(1:ntaxa(blm.rhiz))
com <- veganotu(blm.rhiz)
grps <- vegansam(blm.rhiz)$Crop
grps
library(labdsv)
iva <- indval(com, grps, numitr = 10000)
# Examine the adjusted p-values.
summary(pval.adj2 <- p.adjust(iva$pval, method = "BH"))
(pval.adj2 <- p.adjust(iva$pval, method = "BH"))
source("D:/data/R_default/functions/indval_report.R")
indval.rpt <- indval_report(spe = com, iva = iva, alpha = 0.05, p.adj.method = "BH")
indval.rpt$Family <- tax_table(blm.rhiz)[rownames(indval.rpt), "Family"]
indval.rpt
```

Make a heatmap of these taxa.  

```{r}
temp <- prune_taxa(rownames(indval.rpt), blm.rhiz)
plot_heatmap(temp, taxa.label = "Family", sample.label = "Crop", sample.order = "Crop")

sum(taxa_sums(temp))/sum(taxa_sums(blm.rhiz))

plot_tree(temp, label.tips = "Family", color = "Crop", size = "Abundance")
```

