---
title: "Diversity"
author: "John Quensen"
date: "5/15/2020"
output:
  pdf_document: default
  html_document: default
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, tidy=TRUE, warning=FALSE, message=FALSE, echo=FALSE}
suppressPackageStartupMessages(library(phyloseq))
suppressPackageStartupMessages(library(QsRutils))
suppressPackageStartupMessages(library(ade4))
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(ggordiplots))
suppressPackageStartupMessages(library(hablar))
pretty_int <- function(x) {
  formatC(x, format = "f", big.mark = ",", digits = 0)
}
load("D:/Users/johnq/OneDrive/Kize/phyloseq/expt_sub.rda")
```

1. Does crop rotation increase diversity in relation to maize and soybean mono-culture?  

To compare $\alpha$-diversity measures, there has to be the same number of counts per sample. We begin with expt.sub which has already been filtered of non-bacterial and non-archaeal ASVs and rarefied (sampled) to `r pretty_int(sample_sums(expt.sub)[1])` counts per sample.  

There are several $\alpha$-diversity indices. Some measure richness, some measure evenness, and some combine the two into one measure.  I think the most pertinent for our purposes are the observed number of ASVs as a measure of richness,  chao1 as an estimate of the total number of species (taking into account sampling is incomplete), Pielou's J as a measure of evenness, and Shannon or H' which is related to both richness and evenness.  

First, we have to subset our data to the treatment combinations we want to compare. We begin by sub-setting to harvest time and bulk soil.  

```{r}
bulk.har <- subset_samples(expt.sub, Sampling_time == "Harvest")
bulk.har <- subset_samples(bulk.har, Soil == "Bulk")
bulk.har <- prune_taxa(taxa_sums(bulk.har)>0, bulk.har)
sample_sums(bulk.har)
```

Phyloseq includes wrapper functions for calculating and plotting some $\alpha$-diversities, but not Pielou's J. We can get a quick visual comparison of the other measures among years (seasons) with the plot_richness function.  

```{r}
plot_richness(bulk.har, x = "Season", measures = c("Observed", "Chao1", "Shannon"))
```
Each symbol represents an individual sample. There might be a little more diversity the last year.  

To include Pielou's J, make custom plots, and make statistical comparisons, we need to do some data wrangling.  
```{r}
yr <- vegansam(bulk.har) %>% 
  rownames_to_column("Sample") %>% 
  select(Sample, Season) %>% 
  as_tibble()
yr

temp <-   estimate_richness(bulk.har, measures = c("Observed", "Chao1", "Shannon")) %>% 
  rownames_to_column("Sample") %>% 
  select(-se.chao1) %>% 
  mutate(Pielou = Shannon/log(Observed)) %>% 
  as_tibble()
temp

alpha.div <- 
  full_join(yr, temp, by = "Sample") %>% 
  select(-Sample) %>% 
  pivot_longer(c(Observed, Chao1, Shannon, Pielou), names_to = "Index") %>% 
  dplyr::rename(Diversity = value)
  
alpha.div
```

Make box plots.  

```{r}
ggplot(data = alpha.div, aes(x=Season, y=Diversity)) +
  geom_boxplot() +
  facet_wrap(~Index, scales = "free") 
```

But this is for all three cropping methods pooled. Make separate plots for each season so that we can compare cropping methods for each season/year.  

"Re-make" alpha.div to include both season and crop.  

```{r}
yr <- vegansam(bulk.har) %>% 
  rownames_to_column("Sample") %>% 
  select(Sample, Season, Crop) %>% 
  as_tibble()
yr

alpha.div <- 
  full_join(yr, temp, by = "Sample") %>% 
  select(-Sample) %>% 
  pivot_longer(c(Observed, Chao1, Shannon, Pielou), names_to = "Index") %>% 
  rename(Diversity = value)
  
alpha.div
```

Plot for 2018-2019.  

```{r}
alpha.sub <- dplyr::filter(alpha.div, Season == "2018-2019")

ggplot(data = alpha.sub, aes(x=Crop, y=Diversity)) +
  geom_boxplot() +
  facet_wrap(~Index, scales = "free") +
  ggtitle("2018-2019")
```

Plot for 2017-2018.  

```{r}
alpha.sub <- dplyr::filter(alpha.div, Season == "2017-2018")

ggplot(data = alpha.sub, aes(x=Crop, y=Diversity)) +
  geom_boxplot() +
  facet_wrap(~Index, scales = "free") +
  ggtitle("2017-2018")
```

Plot for 2016-2017.  

```{r}
alpha.sub <- dplyr::filter(alpha.div, Season == "2016-2017")

ggplot(data = alpha.sub, aes(x=Crop, y=Diversity)) +
  geom_boxplot() +
  facet_wrap(~Index, scales = "free") +
  ggtitle("2016-2017")
```

We can also make a plot for all years pooled.  

```{r}
ggplot(data = alpha.div, aes(x=Crop, y=Diversity)) +
  geom_boxplot() +
  facet_wrap(~Index, scales = "free") +
  ggtitle("All seasons/years pooled")
```

Make a statistical test for differences in the observed number of ASVs among crops for the 2018-2019 season. Make a box plot with compact letter displays indicating significant differences among the crops.   

```{r}
# Select season and index.
df <- alpha.div %>% 
  dplyr::filter(Season == "2018-2019") %>% 
  dplyr::filter(Index == "Observed")
df

# Summarize the diversity values.
df_sum <- df %>% 
  group_by(Crop) %>%
  summarize(
    count = n(),
    mean = mean(Diversity, na.rm = TRUE),
    sd = sd(Diversity, na.rm = TRUE),
    median = median(Diversity, na.rm = TRUE),
    IQR = IQR(Diversity, na.rm = TRUE),
    max = max(Diversity, na.rm = TRUE)
  )

df_sum

# Make a non-parametric test for differences among crops.
kruskal.test(Diversity ~ Crop, data = df)

# The result is not quite significant, but contune anyway to show the procedure. Make pairwise tests among the crops.
rslt <- pairwise.wilcox.test(df$Diversity, df$Crop, p.adjust.method = "BH")
rslt

# Show a matrix of the p.values for the pairwise tests.
rslt$p.value

# Get compact letter displays (CLDs) for significant differences if alpha = 0.1.
ltrs<- make_letter_assignments(rslt, significance = 0.1)
cld <- ltrs$Letters
cld

# Get coordinates for the tops of whiskers n the box plot.
box.rslt <- with(df, graphics::boxplot(Diversity ~ Crop, plot = FALSE))
whisker.tops <- box.rslt$stats[5, ]
whisker.tops

# Make a tibble for plotting the CLDs.
x <- c(1:length(cld))
x
df_ltrs <- data.frame(x, whisker.tops, df_sum$max, cld) %>% 
  mutate(max = pmax(whisker.tops, df_sum.max)) %>% 
  select(x, cld, max) %>% 
  as_tibble() %>% 
  hablar::convert(chr(cld))
df_ltrs

# Make the plot.
# Save the plot to a variable.
plt <- ggplot(data = df, aes(x=Crop, y=Diversity)) +
  geom_boxplot() +
  ggtitle("2018-2019")

# Find value for nudging CLDs upward so they do not overlap whisker or maximum value.
lmts <- get_plot_limits(plt)
y.range <- lmts$ymax - lmts$ymin
y.nudge <- 0.05 * y.range

# Make the plot.
plt + geom_text(data = df_ltrs, aes(x=x, y=max, label=cld), nudge_y = y.nudge, size = 5)
```

```{r}
df_ltrs <- function(df = df, grp = "Crop", value = "Diversity", alpha = 0.05) {
  rslt <- pairwise.wilcox.test(pull(df, Diversity), pull(df, Crop))
  ltrs <- make_letter_assignments(rslt, significance = alpha)$Letters
  max.val <- df %>% 
    dplyr::group_by_at(grp) %>% 
    dplyr::summarize_at(.vars = value, .funs = max)
  df_ltrs <- data.frame(max.val, ltrs)
  colnames(df_ltrs) <- c("x", "y", "ltrs")
  return(df_ltrs)
}

df.ltrs <- df_ltrs(df, grp = "Crop", value = "Diversity", alpha = 0.1)
df.ltrs

plt + geom_text(data = df.ltrs, aes(x=x, y=y, label=ltrs), nudge_y = y.nudge, size = 5)

```


```{r}
# Select season and index.
df <- alpha.div %>% 
  dplyr::filter(Season == "2018-2019") %>% 
  dplyr::filter(Index == "Pielou")
df

# Summarize the diversity values.
df_sum <- df %>% 
  group_by(Crop) %>%
  summarize(
    count = n(),
    mean = mean(Diversity, na.rm = TRUE),
    sd = sd(Diversity, na.rm = TRUE),
    median = median(Diversity, na.rm = TRUE),
    IQR = IQR(Diversity, na.rm = TRUE),
    max = max(Diversity, na.rm = TRUE)
  )

df_sum

plt <- ggplot(data = df, aes(x=Crop, y=Diversity)) +
  geom_boxplot() +
  ggtitle("2018-2019")
plt

df.ltrs <- df_ltrs(df, grp = "Crop", value = "Diversity", alpha = 0.1)
df.ltrs

plt + geom_text(data = df.ltrs, aes(x=x, y=y, label=ltrs), nudge_y = y.nudge, size = 5)


```

```{r}
# Make a non-parametric test for differences among crops.
kruskal.test(Diversity ~ Crop, data = df)

# The result is not quite significant, but contune anyway to show the procedure. Make pairwise tests among the crops.
rslt <- pairwise.wilcox.test(df$Diversity, df$Crop, p.adjust.method = "BH")
rslt

# Show a matrix of the p.values for the pairwise tests.
rslt$p.value

# Get compact letter displays (CLDs) for significant differences if alpha = 0.1.
ltrs<- make_letter_assignments(rslt, significance = 0.1)
cld <- ltrs$Letters
cld

# Get coordinates for the tops of whiskers n the box plot.
box.rslt <- with(df, graphics::boxplot(Diversity ~ Crop, plot = FALSE))
whisker.tops <- box.rslt$stats[5, ]
whisker.tops

```

