---
title: "Filter and Rarefy"
author: "John Quensen"
date: "5/13/2020"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r, tidy=TRUE, warning=FALSE, message=FALSE, echo=FALSE}
suppressPackageStartupMessages(library(phyloseq))
suppressPackageStartupMessages(library(QsRutils))
suppressPackageStartupMessages(library(ade4))
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(ggordiplots))
suppressPackageStartupMessages(library(hablar))
pretty_int <- function(x) {
  formatC(x, format = "f", big.mark = ",", digits = 0)
}
# 
# setwd("D:/Users/johnq/OneDrive/Kize/phyloseq")
load("D:/Users/johnq/OneDrive/Kize/phyloseq/expt.rda")
```

There was an NA in the sample data table for one of the replicates. I fixed it in this version of expt.rda.  

Let's check ASVs that should be removed, based on classification. For example, we will remove those classified as uncl_bacteria or uncl_archaea, and those unclassified to a bacterial phylum. First get the names of the taxa ranks, and a list of the unique taxa in the highest rank.  

```{r}
rank_names(expt)
get_taxa_unique(expt, taxonomic.rank = "Domain")
```

Remove ASVs classified as "uncl_domain" and then get a list of phyla.   

```{r}
expt.sub <- subset_taxa(expt, Domain != "uncl_domain")
get_taxa_unique(expt.sub, taxonomic.rank = "Phylum")
```

Remove the phylum "uncl_Bacteria."

```{r}
expt.sub <- subset_taxa(expt.sub, Phylum != "uncl_Bacteria")
```

The phylum "Cyanobacteria/Chloroplast" may contain ASVs classified as chloroplasts. Check if this is so, and remove them if they are present.  

```{r}
cyan <- subset_taxa(expt.sub, Phylum == "Cyanobacteria/Chloroplast")
get_taxa_unique(cyan, taxonomic.rank = "Class")
```

So chloroplasts sequences are present. We need to remove them.  

```{r}
expt.sub <- subset_taxa(expt.sub, Class != "Chloroplast")
```

Re-check that we got rid of all taxa that we wanted to.  

```{r}
get_taxa_unique(expt.sub, taxonomic.rank = "Domain")
get_taxa_unique(expt.sub, taxonomic.rank = "Phylum")
cyan <- subset_taxa(expt.sub, Phylum == "Cyanobacteria/Chloroplast")
get_taxa_unique(cyan, taxonomic.rank = "Class")
```

Now check the number of ASVs per sample.  

```{r}
expt.sub <- prune_taxa(taxa_sums(expt.sub)>0, expt.sub)
summary(sample_sums(expt.sub))
```

Let's "rarefy" our data, i.e. sample to the same number of counts per sample. The minimum is presently `r pretty_int(min(sample_sums(expt.sub)))`. 

```{r}
expt.sub <- rarefy_even_depth(expt.sub, rngseed = 913, replace = FALSE)
expt.sub
save(expt.sub, file = "D:/Users/johnq/OneDrive/Kize/phyloseq/expt_sub.rda")
```

We will use expt.sub for further analyses. It has been filtered of unwanted ASVs and sampled to the same number of counts per sample. I set rngseed to 913 (easily remembered because it is my birthday!) so that the rarefying can be duplicated. Otherwise you would get different results every time you rarefy.  
