---
title: "Phylum Bar Plots"
author: "John Quensen"
date: "5/14/2020"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Begin with the phyloseq object that we filtered and rarefied.  


```{r, tidy=TRUE, warning=FALSE, message=FALSE, echo=FALSE}
suppressPackageStartupMessages(library(phyloseq))
suppressPackageStartupMessages(library(QsRutils))
suppressPackageStartupMessages(library(ade4))
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(ggordiplots))
suppressPackageStartupMessages(library(hablar))
pretty_int <- function(x) {
  formatC(x, format = "f", big.mark = ",", digits = 0)
}
load("D:/Users/johnq/OneDrive/Kize/phyloseq/expt_sub.rda")
```

Examine the levels of sample variables.  

```{r}
sample_variables(expt.sub)
unique(sample_data(expt.sub)[, "Season"])
unique(sample_data(expt.sub)[, "Sampling_time"])
unique(sample_data(expt.sub)[, "Soil"])
unique(sample_data(expt.sub)[, "Crop"])
```

For this example I will make a bar plot of the ten more abundant phyla by crop for the bulk soil at harvest time.  

Subset to bulk soil and harvest time.  

```{r}
p.sub <- subset_samples(expt.sub, Soil == "Bulk")
p.sub <- subset_samples(p.sub, Sampling_time == "Harvest")
p.sub <- prune_taxa(taxa_sums(p.sub) > 0, p.sub)
p.sub
```

# Quick Bar Plot via Phyloseq's plot_bar

Phyloseq includes some functions for analyzing and plotting data. These are quick to run and good for a quick look at your data. For example, we can make a bar plot of the ten more abundant phyla with the following few lines of code.    

```{r}
# Get the names of the 10 more abundant phyla.
p.sub <- tax_glom(p.sub, taxrank = "Phylum")
top10 <- names(sort(taxa_sums(p.sub), decreasing  = TRUE)[1:10])
top10

# Prune the taxa (phyla) to include only these 10 more abundant ones.
p.sub <- prune_taxa(top10, p.sub)

# Make a bar plot.
plot_bar(p.sub, x="Crop", y="Abundance", fill = "Phylum")
```

There are several segments for each of the ten phyla. The lines separate values for individual samples. We can remove the lines by saving the plot and then adding another ggplot command to remove the lines.  

```{r}
plt <- plot_bar(p.sub, x="Crop", y="Abundance", fill = "Phylum")
plt + geom_bar(aes(color=Phylum, fill=Phylum), stat="identity", position="stack")
```

This shows that there are no differences in the relative abundances of the phyla among crops.  

# Customized Bar Plot of Phyla

This plot is sufficient if we are just investigating our data. But if we want a plot for a presentation, we might want more control over how it looks. For example, we could order the phyla by relative abundance instead of alphabetically, with the most abundant on the bottom, and the order in the legend matching the order in the bars.  
  
To do this first convert the OTU table from counts to percentages and check.  

```{r}
p.sub <- transform_sample_counts(p.sub, function(x) 100*(x/sum(x)))
sample_sums(p.sub)
```

Extract the OTU table and change the taxa names to the phyla they represent.  

```{r}
otu <- veganotu(p.sub)
colnames(otu) <- tax_table(p.sub)[colnames(otu), "Phylum"]
colnames(otu)
```

Let's shorten one of the phylum names.  

```{r}
colnames(otu) <- gsub("candidate division ", "", colnames(otu))
colnames(otu)
```

Make otu a tibble.  

```{r}
otu <- otu %>% 
  as.data.frame() %>% 
  rownames_to_column("Sample") %>% 
  as_tibble()
otu
```

Extract the sample data table and convert to a tibble with column names Sample and Crop.  

```{r}
crop <- vegansam(p.sub) %>% 
  rownames_to_column("Sample") %>% 
  as_tibble() %>% 
  select(Sample, Crop)
crop
```

Merge crop and otu by sample names.  

```{r}
df <- full_join(crop, otu, by = "Sample")
df
```

Gather the phyla into a new column named Phylum.  

```{r}
df <- pivot_longer(df, 3:12, names_to = "Phylum")
df
```

Calculate mean abundance by Crop and Phylum.  

```{r}
df_summary <- df %>% 
  group_by(Crop, Phylum) %>% 
  summarize(mean_abundance = mean(value),
            sd_abundance = sd(value))
```

Make Phylum a factor and order the levels by mean abundance. I do this to control the order of the phyla in the plot. Otherwise they will be ordered alphabetically. By setting desc to FALSE, the phylum with the greatest mean abundance will be at the bottom of the bars and legend.    

```{r}
df_summary$Phylum <- fct_reorder(df_summary$Phylum, df_summary$mean_abundance, mean, .desc = FALSE)
df_summary
```

Make a bar plot.

```{r}
ggplot(data = df_summary, aes(x = Crop, y = mean_abundance, fill = Phylum)) +
  geom_col() +
  ylab("Relative Abundance") +
  ggtitle("Top 10 Phyla")
```

The plot can be made in a different way. Unstack the bars so that we can show error bars ( plus and minus the standard deviation) for each phylum mean. Reorder Phylum again so that this time the relative abundances of the phyla decrease from left to right. Use facet_wrap to make separate panels for each crop. Remove X-axis tick marks and tick labels as they are not necessary. 

```{r, eval=FALSE}
df_summary$Phylum <- fct_reorder(df_summary$Phylum, df_summary$mean_abundance, mean, .desc = TRUE)
ggplot(data = df_summary, aes(x = Phylum, y = mean_abundance, fill = Phylum)) +
  geom_col() +
  geom_errorbar(aes(ymin = mean_abundance - sd_abundance,
                    ymax = mean_abundance + sd_abundance), width = 0.5) +
  facet_wrap(~ Crop) +
  theme(axis.text.x=element_blank(),
        axis.ticks.x=element_blank()) +
  ylab("Relative Abundance") +
  ggtitle("Top 10 Phyla")
```

